Empirical Analysis of Performance Bottlenecks in Graph Neural Network Training and Inference with GPUs

### 2020-12-26 NEUROC/sec4-experiments.tex

@@

\paragraph{Comparing Inference and Training}

Figure~\ref{fig:memory_usage_analysis} shows the memory expansion ratios of typical GNNs during inference.
%
Since no intermediate results had to be cached during inference of GGNN, GAT, and GaAN,
the memory expansion ratios of inference were much less than training for the three GNNs (\figurename~\ref{fig:exp_memory_expansion_ratio}).
%
The MERs of inference were 45\% to 83\% (GGNN), 52\% to 61\% (GAT), and 37\% to 69\% (GaAN) of training.
%
However, the MERs were still high, disallowing inferencing with big graphs.
%
To handle big graphs, the sample-based inference was necessary.

@@

%
We also did experiments about the effects of input feature dimension, average degrees, and the number of vertices on GNN inference memory usage.
%
The results were similar to the GNN training phase.
%

@@

With the sampling techniques, GNNs were trained and inferred in a mini-batch manner.
%
Each epoch was decomposed into many batches.
%
In each batch, PyG only sent the sampled subgraph to the GPU to train or inference.
%
The batch size determined the size of the sampled subgraph.
%
Thus, it affected the accuracy, training/inference time, and memory usage of GNNs simultaneously.

@@

\subsubsection{Performance Bottlenecks Analysis}

@@

\paragraph{Training Time}

@@

The sample-based inference was an effective technique to handle big graphs.
%
Taking the node classification task as an example, to predict labels for a given set of vertices $\mathcal{V}_{predict}$, the inference sampler sampled a subgraph containing complete $L$-hop neighborhoods of all vertices in $\mathcal{V}_{predict}$, where $L$ was the number of GNN layers.
%
Since the real-world graphs often had small-world property, the size of the sampled subgraph increased quickly as the number of vertices in $\mathcal{V}_{predict}$ (i.e., batch size) increased.
%
Figure~\ref{fig:sampled_graph_size} shows how the relative batch size ($\frac{|\mathcal{V}_{predict}|}{|\mathcal{V}|}$) affected the average degree and the number of edges of the sampled subgraphs during inference.
%
When the relative batch size was 10\%, the number of edges in the sampled subgraphs were close to the number of edges in the whole graph in most datasets.
%
To limit the memory usage during inference, the batch size used should be very small.

@@

   \caption{Characteristics of the sampled subgraphs produced by the inference sampler under different relative batch sizes.}
   \label{fig:sampled_graph_size}
\end{figure}

However, the overheads brought by sampling and data transferring from CPU to GPU became obvious when the batch size was small.
%
Figure~\ref{fig:time_breakdown_of_inference_sampler} shows the time breakdowns of sample-based inference on different datasets.
%
On the \texttt{amp} and \texttt{cam} datasets, the sampling time even accounted for near half of the total inference time.
%
The results indicated that the current implementation of inference sampler in PyG was also inefficient.
%
Improving its efficiency could significantly reduce the inference time.

@@

Since the size of the sampled subgraph was much smaller than the original graph, the accuracy of the GNN models got by sampling techniques might be different from the full-batch training.
%
To find out how the batch size affected the test accuracy, we trained GNNs with different batch sizes.
%
\figurename~\ref{fig:exp_sampling_relative_batch_size_accuracy_graphsage} and \figurename~\ref{fig:exp_sampling_relative_batch_size_accuracy_cluster} compare the test accuracy achieved by sampling techniques with the test accuracy of the full-batch training.
%
For each combination of dataset and GNN, we chose the hyper-parameters that achieved the highest accuracy according to the experimental results in Section 4.1.4.

@@

\paragraph{Summary of Sampling Techniques}
% 
The sampled subgraphs had lower average degrees than the original graph.
%
With small batch sizes, the sampling techniques could significantly reduce the training/inference time per batch and the peak memory usage.
%
However, small batch sizes could not make full use of the computing power of a GPU.
%
With big batch sizes, the current implementation of the sampling techniques in PyG was inefficient.
%
The time spent on the sampling phase and the data transferring phase even exceeded the training/inference phase.

>>>>

\subsubsection{Effects on Accuracy}
\label{sec:hyper_parameter_accuracy}

@@

\paragraph{Inference}

Figure~\ref{fig:memory_usage_analysis} shows the memory expansion ratios of typical GNNs during inference.
%
Since no intermediate results had to be cached during inference of GGNN, GAT, and GaAN,
the MERs of inference were much less than training for the three GNNs (compared with \figurename~\ref{fig:exp_memory_expansion_ratio}).
%
The MERs of inference were 45\% to 83\% (GGNN), 52\% to 61\% (GAT), and 37\% to 69\% (GaAN) of training.
%
However, the MERs were still high, disallowing inferencing with big graphs.
%
To handle big graphs, the sample-based inference was necessary.

@@

%
We also conducted experiments about the effects of input feature dimensions, average degrees, and the number of vertices on the memory usage of inference.
%
The results were similar to GNN training.
%

@@

With the sampling techniques, GNNs were trained and inferred in a mini-batch manner.
%
Each epoch was decomposed into many batches.
%
In each batch, PyG only sent the sampled subgraph to the GPU to train or inference.
%
The batch size determined the size of the sampled subgraph.
%
Thus, it affected the accuracy, training/inference time, and memory usage simultaneously.

@@

\subsubsection{Performance Bottlenecks in Sample-based Training}

@@

@@

The sample-based inference was an effective technique to conduct inference on big graphs.
%
Taking the node classification task as an example, to predict labels for a given set of vertices $\mathcal{V}_{predict}$, the inference sampler sampled a subgraph containing complete $L$-hop neighborhoods of all vertices in $\mathcal{V}_{predict}$, where $L$ was the number of GNN layers.
%
Since the real-world graphs often had small-world property, the size of the sampled subgraph increased quickly as the number of vertices in $\mathcal{V}_{predict}$ (i.e., batch size) increased.
%
Figure~\ref{fig:sampled_graph_size} shows how the relative batch size ($\frac{|\mathcal{V}_{predict}|}{|\mathcal{V}|}$) affected the average degree and the number of edges of the sampled subgraphs during inference.
%
When the relative batch size was 10\%, the number of edges in the sampled subgraphs were close to the number of edges in the whole graph in most datasets.
%
To limit the memory usage during inference, the batch size used should be very small in inference.

@@

   \caption{Sizes of sampled subgraphs produced by the inference sampler under different relative batch sizes.}
   \label{fig:sampled_graph_size}
\end{figure}

However, the overheads brought by sampling and data transferring from CPU to GPU became obvious when the batch size was small.
%
Figure~\ref{fig:time_breakdown_of_inference_sampler} shows the time breakdowns of sample-based inference on different datasets.
%
On the \texttt{amp} and \texttt{cam} datasets, the sampling time even accounted for near half of the total inference time.
%
The results indicated that the current implementation of inference sampler in PyG was also inefficient.
%
Improving its efficiency could significantly reduce the sample-based inference time.

@@

Since the size of the sampled subgraph was much smaller than the original graph, the accuracy of the GNN models trained with sampling techniques might be different from the full-batch training.
%
To find out how the batch size affected the test accuracy, we trained GNNs with different batch sizes.
%
\figurename~\ref{fig:exp_sampling_relative_batch_size_accuracy_graphsage} and \figurename~\ref{fig:exp_sampling_relative_batch_size_accuracy_cluster} compare the test accuracy achieved by sampling techniques with the test accuracy of the full-batch training.
%
For each combination of dataset and GNN, we chose the hyper-parameters that achieved the highest accuracy according to the experimental results in Section~\ref{sec:hyper_parameter_accuracy}.

@@

\paragraph{Summary of Sampling Techniques}
% 
The sampled subgraphs produced by the neighbor sampler and the cluster sampler had lower average degrees than the original graph.
%
With small batch sizes, the sampling techniques could significantly reduce the training/inference time per batch and the peak memory usage.
%
However, the batch size should be small to limit the memory usage, especially for inference.
%
The current implementation of the sampling techniques in PyG was inefficient to handle small batches.
%
Small batches could not make full use of the computing power of a GPU.
%
The time spent on the sampling phase and the data transferring phase could even exceed the training/inference phase.
